%!TEX root = ../main.tex


\section{Learning Objectives}
\label{section:learning-objectives}

In the FACT-AI course, we aim to make students aware of two types of responsibility: (i) towards society in terms of potential implications of their research, and (ii) towards the research community in terms of producing reproducible research. 
In this section, we outline the learning objectives for the FACT-AI course and explain how it fits within the context of the MSc AI program at \OurUniversity{}. 

\begin{table}[t]
\caption{The first year of the MSc AI program at \OurUniversity{}, 2019--2020.}
\label{tab:msc_program}
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l c@{~}c@{~}c c@{~}c@{~}c c}
\toprule
Course & \multicolumn{3}{c}{Sem.\ 1} & \multicolumn{3}{c}{Sem.\ 2} & EC \\
\midrule
     Computer Vision 1 & $\blacksquare$ & $\square$ & $\square$ & $\square$ & $\square$ & $\square$ & 6 \\
     Machine Learning 1 & $\blacksquare$ & $\square$ & $\square$ & $\square$ & $\square$ & $\square$ & 6 \\
     Natural Language Processing 1 & $\square$ & $\blacksquare$ & $\square$ & $\square$ & $\square$ & $\square$ & 6 \\
     Deep Learning 1 & $\square$ & $\blacksquare$ & $\square$ & $\square$ & $\square$ & $\square$ & 6\\
     Fairness, Accountability, Confidentiality  & $\square$ & $\square$ & $\blacksquare$ & $\square$ & $\square$ & $\square$ & 6 \\
     and Transparency in AI \\
     Information Retrieval 1 & $\square$ & $\square$ & $\square$ & $\blacksquare$ & $\square$ & $\square$ & 6 \\
     Knowledge Representation and Reasoning & $\square$ & $\square$ & $\square$ & $\blacksquare$ & $\square$ & $\square$ & 6 \\
     Elective 1 & $\square$ & $\square$ & $\square$ & $\square$ & $\blacksquare$ & $\square$ & 6 \\
     Elective 2 & $\square$ & $\square$ & $\square$ & $\square$ & $\blacksquare$ & $\square$ & 6 \\
     Elective 3 & $\square$ & $\square$ & $\square$ & $\square$ & $\square$ & $\blacksquare$ & 6 \\
\bottomrule
\end{tabular}
\end{table}


Table~\ref{tab:msc_program} shows the setup of the first year of the 2-year MSc AI program. Each semester at \OurUniversity{} is divided into three periods: two 8-week periods followed by one 4-week period. During an 8-week period, students follow two courses in parallel. During the 4-week period, they only follow a single course. 
The FACT-AI course takes place during the 4-week period at the end of the first semester, after students have taken Computer Vision 1, Machine Learning 1, Natural Language Processing 1 and Deep Learning 1. 
It is the only course students follow during this period, so we believe it is beneficial to have them focus on one main project -- reproducing an existing FACT-AI paper. 
The learning objectives for the course are as follows: 

\begin{itemize}
	\item \textbf{LO \#1: Understanding FACT topics.} Students can explain the major notions of fairness, accountability, confidentiality, and transparency that have been proposed in the literature, along with their strengths and weaknesses.
	\item \textbf{LO \#2: Understanding algorithmic harm.} Students can explain, motivate, and distinguish the main types of algorithmic harm, both in general and in terms of concrete examples where AI is being applied.
	\item \textbf{LO \#3: Familiarity with FACT methods.} Students are familiar with recent peer-reviewed algorithmic approaches to fairness, accountability, confidentiality, and transparency in the literature. 
    \item \textbf{LO \#4: Reproducing FACT solutions.} Students can assess the degree to which recent algorithmic solutions are effective, especially with respect to the claims made in the original papers, while understanding their limitations and shortcomings. 
\end{itemize}
