% !TEX root = thesis-main.tex
{\pagestyle{empty}

{
\begin{center}

\noindent
\textbf{Acknowlegements} \\ \vskip .5cm
\end{center}
}





\noindent
\if
Moving to a foreign country to do a PhD on a foreign topic was not something I had originally planned, but it has nonetheless turned out to be a really rewarding experience. 
This thesis is the culmination of that experience, and there are many people without whom it would not have been possible. 

First and foremost, I want to thank my supervisors who guided me through this whole process. 
Maarten -- thank you for always listening to me and advocating for me.
Hinda -- thank you for your unwavering support and for always being on my team. 

Next, I'd like to thank the committee members who reviewed my thesis. 
Miroslav -- thank you for motivating me to pursue a math major during my undergrad. This decision has fueled everything I've done since then and I'm not sure I would have been confident enough to do it without your encouragement.
%
Fernando -- thank you for taking over the FACT-AI course we created for the MSc AI program. I'm really looking forward to seeing how it develops over the next few years. 
%
Clarisa -- thank you for giving me advice about the challenges I encountered during my fellowship project. Your guidance helped me prioritize and carve out the next steps. 
%
Fabrizio -- thank you for letting me do my internship project unofficially after I was unable to do a formal internship due to the pandemic. This project is my favorite chapter in this thesis and it has inspired me to shift my research into a new direction. 
%
Max -- thank you for allowing me to pursue this new direction under your guidance. 





I want to thank my paranymphs for standing with me as I defend my thesis. 
Maartje and Maurits -- thank you for being my main sources of support within our group and for sharing all of the ups and downs of doing a PhD with me. I also want to thank the many great people I've met while working at the University of Amsterdam:
Alexey, 
Ali, 
Ali, 
Amir, 
Antonis, 
Arezoo, 
Artem, 
Boris, 
Chang, 
Christof, 
Chuan,
Dan,  
David, 
Gabriel, 
Hamid, 
Ilya, 
Jiahuan, 
Jie, 
Jin, 
Julia, 
Julien, 
Katya, 
Ke, 
Maria, 
Marlies, 
Mohammad, 
Mostafa, 
Mozdheh, 
Olivier, 
Pengjie, 
Petra, 
Sam, 
Sami, 
Shaojie,
Spyretta, 
Svitlana, 
Thomas, 
Trond, 
Vera, 
Yangjun, and 
Yifan -- thank you for making Science Park (and especially our mint green container) a fun place to work.  
%
In particular, I want to thank 
Harrie,
Hosein, 
Nikos, 
Rolf, 
Christophe, 
Bob, and 
Tom for your leadership in the group, and of course for all the borrels, biertjes, and bitterballen. 
Marzieh, 
Mariya, and
Mahsa
 -- thank you for your honesty, friendship, and support over these last few years. 
Ziming -- thank you for letting me babysit your cat for a summer, it was truly one of the highlights of my PhD. 
I also want to thank the many bright students I had: Michael, Stefan, Kim, Puja, Fije -- thank you for teaching me way more than I taught you. 




During my PhD, I had the opportunity to do a fellowship at the Partnership on AI. Rebecca -- thank you for leading me through my time at the Partnership and for your trust in me amid all the twists and turns involved in doing real-world research. 
\pagebreak


I've had three lovely roommates while living in Amsterdam, each of which were living with me at very different phases of my PhD. 
Elisabet -- thank you for helping me through the challenges and uncertainties of early PhD life and for showing me every single specialty coffee shop in Amsterdam. 
%
Maria -- thank you for quarantining with me and supporting me through hairdresser fiascos and global pandemics (not sure what's worse!). Luckily we can always fall back on falafel salads and eggless desserts. 
%
Camila -- thank you for supporting my leopard print addiction and for always being up for something fun. 
%
Special shoutout to my Kelowna roommates Danielle and Bree, who only lived with me briefly during my PhD, but have nonetheless contributed to many ``good vibes'' and sometimes questionable memories. \\ 


I've learned that having a support system outside the research world is incredibly important while pursuing a PhD. 
Sophie, Tamara, Juliette, Jago, and Thomas -- thank you for adopting me at Mystic Garden all those years ago and for the many parties, picnics, pizzas, and pals we've shared since then.
%
Duffey, Allie, Kristine, Craig, Devra, Jeff, and Aly -- thank you for not forgetting about me even though I only come home once a year. 
%
Thomas -- thank you for showing me how to be a kinder, softer person and for always listening to the many things I have to say (ook in het Nederlands). You are what I look forward to. 
Bedankt aan de hele familie Jak om me welkom te heten met open armen en open harten. \\ 


Naravno se moram zahvaliti na\v{s}om rajom. 
Hvala mojim Para\'{c}incima: Deda Uj\v{c}e, Baka Ujna, Ana, Sale, Sanja, Mi\v{s}o, Margarita, Marija i Nikola, i mojim drugim roditeljim iz Kanade: Ksenija, Mi\v{s}o, Srdjan, Ksenija, Stanko, \v{Z}eki I Gogo. Vi ste mi uvijek u srcu.
%
Najva\v{z}nije, ho\'{c}u da se zahvalim svojim roditeljima, Vesna i Tiho, koji su meni sve mogu\'{c}e u\v{c}inili. Hvala mojoj Baki koja me je ``dva puta rodila'' i cuvala od malenog. Ja vas obozavam, i ovo je za vas. \\ \vskip .3cm

\fi


\RaggedLeft{
Gabriel Bénédict \\
Amsterdam \\
January 2024}



\clearpage
}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "thesis-main"
%%% End:
